---
title: "MAGLi_cCDP_MCPModGeneral_package"
format: html
editor: visual
---

# MAGL inhibitor Ph2 design

```{r}
# install.packages("MCPModGeneral")

require(DoseFinding)
require(MCPModGeneral)

```

```{r}

# candidate set of models for the mean response on the logit scale:

logit <- function(p) log(p / (1 - p))
inv_logit <- function(y) 1 / (1 + exp(-y))

doses <- c(0, 1, 2, 3)/3 # normalized equidistant doses
# doses <- c(0, 1, 2, 3,4,5)/5 # normalized equidistant doses

p.ocr <- 0.43
p.magli <- p.ocr * 0.5

mods5 <- Mods(emax = c(0.25, 0.5), 
             sigEmax = rbind(c(0.25, 3), c(0.5, 4)), 
             betaMod = c(1.5, 0.1),
             placEff = logit(p.ocr), 
             maxEff = logit(p.magli) - logit(p.ocr), 
             doses = doses)


# plotMods(mods5) ## plot candidate models on logit scale

plotMods(mods5, trafo = inv_logit) ## plot candidate models on probability scale

# plot(mods5, plotTD = TRUE, Delta = -0.1)

```

```{r}
# power calculation for dose-response (PoC)

powMCTGen(nSample = 81, 
          doses = doses, family = "binomial", link = "logit",
          Ntype = "arm", altModels = mods5, 
          alpha = 0.025)

# min power 80%

Assuming 10% DO rate 81/0.9 = 90 per arm
```

# Estimating the doses of interest

## Precision around the Target Dose (TD)

Precision around the TD varies with the assumed dose-response model.

```{r}
Smat <- planModPrepare(nSample = 81, family = 'binomial', link = 'log', 
               theoResp = getResp(mods5)[,3],
               doses = doses,
               Ntype = "arm")

pObj1 <- planMod(model = "emax",
        altModels = mods5,
        S = Smat,   # Either a vector ‘⁠n⁠’ and ‘⁠sigma⁠’ or ‘⁠S⁠’ need to be specified.
        simulation = T,
        nSim = 10000, 
        cores = 10)

summary(pObj1, Delta = -(logit(p.magli) - logit(p.ocr))*0.9)

```

## Probability of estimating ED90 within certain bounds.

What is the probability that the placebo-corrected response associated with the estimated dose will be within 75% to 99% of the maximum response attainable within the dose range used.

```{r}

# Compute covariance matrix
Smat <- planModPrepare(nSample = 81, family = 'binomial', link = 'log', 
               theoResp = getResp(mods5)[ , 3],
               doses = doses,
               Ntype = "arm")

pObj2 <- planMod(model = "emax",
        altModels = mods5,
        S = Smat,   # Either a vector ‘⁠n⁠’ and ‘⁠sigma⁠’ or ‘⁠S⁠’ need to be specified.
        simulation = T,
        p = 0.6,
        pLB = 0.4, pUB = 0.8,
        nSim = 10000, 
        cores = 10)

print(pObj2)
```

For example, P(ED0.9) gives the probability that the estimated ED0.6 if within the true ED0.40 and ED0.80 .

```{r}
sessionInfo()
```
