---
title: "MAGLi Biomarkers Sample Size"
author: "Shamil Sadikhov"
format: html
editor: visual
---


```{r}
#| echo: false
require(MCPMod)
require(ggplot2)
require(DoseFinding)
require(pwr)

```

## 

# CSF Osteopontin

The primary endpoint, CSF osteopontin, was significantly reduced from the mean baseline value 322 ng/mL (95% CI 257--387) by a mean of 65 ng/mL (95% CI 34--96; figure 2) at week 60

Publication: https://n.neurology.org/content/neurology/82/17/1499.full.pdf



# CDP designs for eSPC 
old link: https://docs.google.com/presentation/d/1v0OEwxsEp_Bn-t-4LX9eDWQHYS_vPv8T59C0kUZmuQg/edit?usp=sharing


```{r}
#| echo: false

# CSF Biomarker data

## NFL
nfl.bl <- 1160.30
nfl.52w.mean <- 809.69
nfl.52w.chg <- -261.85
nfl.52w.chg.sd <- 1175.72
nfl.nsize <- 28

## YKL-40
ykl40.bl <- 219.59
ykl40.w52 <- 203.44
ykl40.chg.w52 <- -20.86
ykl40.chg.sd <- 37.43
ykl40.nsize <- 27
```

```{r}

```

# MCP-Mod

```{r}
#| echo: false
#| eval: false
#| 
## variance from the natalizumab osteopontin change from baseline

require(MCPMod)

var <- sd.chg^2

# doses <- c(0, 0.25, 0.5, 0.75, 1) 
doses1 <- c(0, 0.33, 0.67, 1)



ed50 <- guesst(d = c(0.5, 0.75), p = c(0.7, 0.9), model = "emax") 

mods1 <- list(emax = c(ed50), sigEmax = c(ed50, 0.9), betaMod = c(0.5, 1),

logistic = matrix(c(0.25, 0.7, 0.09, 0.06), byrow = FALSE, nrow = 2))

mods1 <- list( emax = c(ed50), sigEmax = c(ed50, 0.9)) 
mods2 <- list( emax = c(ed50), sigEmax = c(ed50, 5), exponential = c(0.6))

fMod <- fullMod(models = mods2, doses = doses1, base = 322, maxEff = -65, scal = 200) 

plot(fMod) # automatically calls the plotModels function

# scal - Scale parameter for the beta model (default: 20 percent larger than maximum dose)

plotModels(mods1, doses, base = 322, maxEff = -70, scal = 1) 
plotModels(mods1, doses1, base = 322, maxEff = -70, scal = 1)


fmods1 <- fullMod(mods2, doses1, base = 322, maxEff = -70, scal = 1)

# Power

# sigma - Expected standard deviation

pM <- powerMM(fmods1, sigma = sqrt(var), alpha = 0.05, lower = 5, upper = 40, step = 2)


```

```{r}
plot(pM, line.at = 0.9, models = "none") 
plot(pM, line.at = 0.8, models = "none")
```



# New request Dec 5 2023

```{r}
#| echo: false


doses1 <- c(0, 1, 2, 3)/3

# YKL-40 Glycoprotein in PPMS (n=31 OBOE)

# change at w52
ykl40.bl <- 219.59
ykl40.w52 <- 203.44
ykl40.chg.w52 <- -20.86
ykl40.chg.sd <- 37.43

magli.chg.w52 <- 1.5 * ykl40.chg.w52 # 50% more relative reduction

mods1 <- DoseFinding::Mods(emax = c(0.25, 0.5), 
             sigEmax = rbind(c(0.25, 3), c(0.5, 4)), 
             betaMod = c(1.5, 0.1),
             placEff = ykl40.chg.w52, 
             maxEff = magli.chg.w52 - ykl40.chg.w52, 
             doses = doses,
             direction = "decreasing")


```

```{r}
plotMods(mods1)
```


```{r}
#| echo: false


## Optimal Contrasts
## here we have balanced sample sizes across groups, so we select w = 1
## otherwise would select w proportional to group sample sizes
contMat <- optContr(mods1, w=1)
print(contMat)
plot(contMat)


pM <- powerMM(fmods1, sigma = ykl40.chg.sd,  alpha = 0.05, lower = 5, upper = 40, step = 2)

pM <- powMCT(contMat, altModels = mods1, n = 200, alpha = 0.025, sigma = ykl40.chg.sd)
print(pM)

sampSizeMCT(upperN = 200, contMat = contMat, altModels = mods1, sigma = ykl40.chg.sd,
            power = 0.8, alRatio = rep(1, length(doses)), alpha = 0.025, Ntype = "arm")

```




```{r}
# YKL40
# power using pairwise comparisons dunnet test



```

# CDP option A1 ‘base case’ with signal seeking Ph2a (BMs) 
 

```{r}
#| echo: false
#| eval: false


# Ph2a design, add on to SoC
# 2 arms

# require(pwr)
# 
# 
# ## YKL-40
# ykl40.bl <- 219.59
# ykl40.w52 <- 203.44
# ykl40.chg.w52 <- -20.86
# ykl40.chg.sd <- 37.43
# ykl40.nsize <- 27
# 
# 
# # YKL-40
# # 50% relative reduction in levels relative to ocrevus
# d.ykl40 <- ykl40.chg.w52*0.5/(sqrt(2)*ykl40.chg.sd)
# ykl40.pwr <- pwr.t.test(d = d.ykl40, power = 0.8, sig.level = 0.1, type = "two.sample", alternative = "two.sided")
# 
# plot(ykl40.pwr)

```
```{r}
# CSF YKL-40 still
require(prwss)

# 50% relative reduction in CSF YKL-40 concentration change from bl relative to control
ykl40.pwrres <- pwrss.t.2means(mu1 = ykl40.chg.w52, mu2 = ykl40.chg.w52 *0.5, sd1 = ykl40.chg.sd, sd2 = ykl40.chg.sd, kappa = 1, 
               power = 0.8, alpha = 0.1,
               alternative = "not equal")

```

```{r}
# NFL


## NFL
nfl.bl <- 1160.30
nfl.52w.mean <- 809.69
nfl.52w.chg <- -261.85
nfl.52w.chg.sd <- 1175.72
nfl.nsize <- 28

# CSF Nfl
# 50% relative reduction in levels relative to ocrevus

# d - difference between the means divided by the pooled standard deviation

# d.nfl <- nfl.52w.chg*0.5 / (sqrt(2) * nfl.52w.chg.sd)
# 
# nfl.pwr <- pwr.t.test(d = d.nfl, power = 0.8, sig.level = 0.1, type = "two.sample", alternative = "two.sided")
# 
# plot(nfl.pwr)

```

```{r}
# CSF NFL still
require(prwss)

# 50% relative reduction in CSF NFL concentration change from bl relative to control
nfl.pwrres <- pwrss.t.2means(mu1 = nfl.52w.chg  , mu2 = nfl.52w.chg *0.5, sd1 = nfl.52w.chg.sd, sd2 = nfl.52w.chg.sd, kappa = 1, 
               power = 0.8, alpha = 0.1,
               alternative = "not equal")

```
```{r}
# MRI T1 non-enhancing volume change form baseline

t1vol.w48.chg <- 0.406 
t1vol.w48.sd <- 1.81

# complete reduction in vol relative to ocrevus

# d - difference between the means divided by the pooled standard deviation

# d.t1v <- (0-t1vol.w48.chg)/t1vol.w48.sd
# 
# t1v.pwr <- pwr.t.test(d = d.t1v, power = 0.8, sig.level = 0.1, type = "two.sample", alternative = "two.sided")
# 
# plot(nfl.pwr)

```

```{r}
# T1 lesion volume still
require(prwss)

# 50% relative reduction in T1 volume change from bl relative to control
t1vol.pwrres <- pwrss.t.2means(mu1 = t1vol.w48.chg , mu2 = t1vol.w48.chg*0.5, sd1 = t1vol.w48.sd, sd2 = t1vol.w48.sd, kappa = 1, 
               power = 0.8, alpha = 0.1,
               alternative = "not equal")

```


# MCP-Mod
Sample size to reduce T1 non-enhancing volume change to 0 at w48

```{r}


## Optimal Contrasts
## here we have balanced sample sizes across groups, so we select w = 1
## otherwise would select w proportional to group sample sizes
contMat <- optContr(mods1, w=1)
print(contMat)
plot(contMat)

mods_t1 <- DoseFinding::Mods(emax = c(0.25, 0.5), 
             sigEmax = rbind(c(0.25, 3), c(0.5, 4)), 
             betaMod = c(1.5, 0.1),
             placEff = t1vol.w48.chg, 
             maxEff = 0 - t1vol.w48.chg, 
             doses = doses,
             direction = "decreasing")

pM <- powMCT(contMat, altModels = mods_t1, n = 200, alpha = 0.025, sigma = ykl40.chg.sd)
print(pM)

sampSizeMCT(upperN = 200, contMat = contMat, altModels = mods_t1, sigma = t1vol.w48.sd,
            power = 0.8, alRatio = rep(1, length(doses)), alpha = 0.025, Ntype = "arm")

```

# CDP option A2 - ‘base case’ with dose finding on BMs in Ph2a 

```{r}

# Ph2a design, add on to SoC
# 4 arms

```
